Windowed Residual-Minimization Model Reduction for Dynamical Systems

### 2019-09-19 section4/spatial_numerical.tex
latest

where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. \methodAcronym\ solves the minimization problem over each time window, 

@@

&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }

@@

& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =

@@

@@

& \genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}  = \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{arg\,min } }

@@

\genstateDiscreteArg{n,0} =

@@

@@

\caption{\spatialAcronym\ trial space: algorithm for the direct solution technique with the Gauss--Newton method and a linear multistep method over the $n$th window}

@@

\begin{algorithm} \caption{\spatialAcronym\ trial space: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $tol$; damping factor, $\rho$; growth factor, $\fbsmGrowth$; decay
factor, $\fbsmDecay$} 
\Output{Stationary point, $\genstateArgnt{n}$ }
\textbf{Online Steps:}\\ 
$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
guess for costate}\\ 
$\genstateArgnt{n}_1 \solveAlgo \basisspace^T \stweightingMatArg{n}
\basisspace \genstateDotArgnt{n}_1  -  \basisspace^T \stweightingMatArg{n}
\velocity(\veloargsromntArg{1}) =  \controllerArgnt{n}$ 
% Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
% obtain $\genstate_0^{n}(t)$ 
\Comment{Solve~\eqref{eq:lspg_continuous} with initial guess to
controller}\\ 
%Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
%$\objectiveArg{n}({\genstate_1^n(t)})$
%\Comment{Evaluate objective function}\\
$i \leftarrow 1$ \Comment{Set counter}\\
\While{$tol \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
\small{
\begin{multline*}
\adjointArgnt{n} \underset{\text{solve}}{\leftarrow} 
\frac{d}{dt} \adjoint^n  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \basisspace \mass^{-1} \adjoint^n = \\ -\bigg[\basisspace^T \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspace \mass^{-1} \basisspace^T \bigg)  \stweightingMatArg{n} \bigg( \basisspace \dot{\genstate}_i^n   -   \velocity( \basisspace \genstate_i^n  +\stateIntercept,t) \bigg) \bigg] 
\end{multline*} }
\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate state} \\
%Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
$\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
$i \leftarrow i+1$ \Comment{Update Counter}\\
$\genstateArgnt{n}_i \solveAlgo \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArgnt{n}_i   -  \basisspace^T \stweightingMatArg{n} \velocity(\basisspace \genstate^n_i + \stateIntercept,t) =  \controllerArgnt{n} $
\Comment{Solve~\ref{eq:lspg_continuous} with updated guess to costate}\\
%$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
\uIf{ $\objectiveArg{n}({\genstate_i^n(t)}) \le \objectiveArg{n}({\genstate_{i-1}^n(t)})$}
{
$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the damping factor}\\
}
\Else{
$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the damping factor}\\ 
$\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
}
}
Return converged solution, $\genstateArgnt{n}= \genstateArgnt{n}_i$
\end{algorithm}
\subsubsection{Numerical Considerations for the Forward and Backwards Problems}
The FBSM requires solving~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
The numerical implementation of the FBSM requires two main ingredients: (1.) temporal discretization schemes for the forward and backward problems and (2.) 
an efficient method for computing action of the Jacobian transpose on a vector.

>>>>

where $\quadWeightsLMSScalarArg{n}{i} \in \RR{}$ are quadrature weights. 
Over each window, \methodAcronym\ computes the generalized coordinates $\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}}$, for $n=1,\ldots,\nslabs$, as the solution to the minimization problems,

@@

&\underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{minimize } }

@@

& \text{subject to} \hspace{0.2 in}  \stateyDiscreteArg{n,0} =

@@

%\methodAcronym\ solves the minimization problem over each time window, 
%&\approxstateDiscreteArg{n,0},\ldots,\approxstateDiscreteArg{n,\nstepsArg{n}} = \underset{\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}} \in \trialspace}{\text{arg\,min } }
%\objectiveArgLMS{n} (\stateyDiscreteArgnt{n,0},\ldots,\stateyDiscreteArgnt{n,\nstepsArg{n}}) \\
%& \text{subject to} \hspace{0.2 in}  \approxstateDiscreteArg{n,0} =
%\begin{cases} \approxstateDiscreteArg{n-1,\nstepsArg{n-1}} & n = 2,\ldots,\nslabs,\\
%\approxstateIC & n=1. \end{cases} \end{align*}

@@

& \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{minimize } }

@@

\genstateyDiscreteArg{n,0} =

@@

%\begin{equation}\label{eq:obj_gen_lms_final}
%\begin{split}
%& \genstateDiscreteArg{n,0},\ldots,\genstateDiscreteArg{n,\nstepsArg{n}}  = \underset{\genstateyDiscreteArg{n,0},\ldots,\genstateyDiscreteArg{n,\nstepsArg{n}}}{\text{arg\,min } }
%\objectiveArgLMS{n} (\basisspace \genstateyDiscreteArg{n,0} + \stateIntercept,\ldots,\basisspace \genstateyDiscreteArg{n,\nstepsArg{n}} + \stateIntercept) \\ 
%& \text{subject to} \hspace{0.2 in}
%\genstateDiscreteArg{n,0} =
%\begin{cases} \genstateDiscreteArg{n-1, \nstepsArg{n-1}} & n = 2,\ldots,\nslabs \\
%\basisspace^T(\stateFOMIC - \stateIntercept)& n=1. \end{cases} 
%\end{split}

@@

\caption{\spatialAcronym\ trial subspace: algorithm for the direct solution technique with the Gauss--Newton method and a linear multistep method over the $n$th window}

@@

\begin{algorithm} \caption{\spatialAcronym\ trial subspace: algorithm for the FBSM over the $n$th window.} \label{alg:st_iter} 
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{tolerance, $tol$; damping factor, $\rho$; growth factor, $\fbsmGrowth$; decay
factor, $\fbsmDecay$} 
\Output{Stationary point, $\genstateArgnt{n}$ }
\textbf{Online Steps:}\\ 
$\genstate^n_0 \leftarrow \bz$ \Comment{Set initial guess for state} \\
$\controllerArgnt{n} \leftarrow \boldsymbol 0$ \Comment{Set initial
guess for costate}\\ 
$\genstateArgnt{n}_1 \solveAlgo \basisspace^T \stweightingMatArg{n}
\basisspace \genstateDotArgnt{n}_1  -  \basisspace^T \stweightingMatArg{n}
\velocity(\veloargsromntArg{1}) =  \controllerArgnt{n}$ 
% Eq.~\ref{eq:lspg_continuous} with $\adjoint^n(t) = \controllerArg{n}{t}$ to
% obtain $\genstate_0^{n}(t)$ 
\Comment{Solve~\eqref{eq:lspg_continuous} with initial guess to
controller}\\ 
%Set: $\genstate^n_1 = \genstate$ \Comment{Set state} \\
%$\objectiveArg{n}({\genstate_1^n(t)})$
%\Comment{Evaluate objective function}\\
$i \leftarrow 1$ \Comment{Set counter}\\
\While{$tol \le \int_{\timeStartArg{n}}^{\timeEndArg{n}} \norm{\genstate^n_{i}(t) - \genstate^n_{i-1}(t) }dt $}{
\small{
\begin{multline*}
\adjointArgnt{n} \underset{\text{solve}}{\leftarrow} 
\frac{d}{dt} \adjoint^n  + \basisspace^T \bigg[\frac{\partial \velocity}{\partial \stateyDiscrete}(\basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \basisspace \mass^{-1} \adjoint^n = \\ -\bigg[\basisspace^T \bigg[ \frac{\partial \velocity}{\partial \stateyDiscrete} ( \basisspace \genstate^n_i + \stateIntercept,t) \bigg]^T \stweightingMatArg{n} \bigg( \mathbf{I} -   \basisspace \mass^{-1} \basisspace^T \bigg)  \stweightingMatArg{n} \bigg( \basisspace \dot{\genstate}_i^n   -   \velocity( \basisspace \genstate_i^n  +\stateIntercept,t) \bigg) \bigg] 
\end{multline*} }
\Comment{Solve~\eqref{eq:lspg_adjoint} to obtain guess to costate state} \\
%Solve Eq.~\ref{eq:lspg_adjoint} with $\genstate^n(t) = \genstate^n_i(t)$ to obtain $\adjoint^n(t)$ \\
$\controllerArgnt{n}  \leftarrow \rho \controllerArgnt{n} + (1 - \rho) \adjointArgnt{n}$ \Comment{Weighted update to costate}\\
$i \leftarrow i+1$ \Comment{Update Counter}\\
$\genstateArgnt{n}_i \solveAlgo \basisspace^T \stweightingMatArg{n} \basisspace \genstateDotArgnt{n}_i   -  \basisspace^T \stweightingMatArg{n} \velocity(\basisspace \genstate^n_i + \stateIntercept,t) =  \controllerArgnt{n} $
\Comment{Solve~\ref{eq:lspg_continuous} with updated guess to costate}\\
%$\objectiveArg{n}({\genstate_i^n(t)})$ \Comment{Evaluate objective function}\\
\uIf{ $\objectiveArg{n}({\genstate_i^n(t)}) \le \objectiveArg{n}({\genstate_{i-1}^n(t)})$}
{
$\rho \leftarrow \text{min}(\rho \fbsmGrowth,1)$ \Comment{Grow the damping factor}\\
}
\Else{
$\rho \leftarrow \frac{\rho }{ \fbsmDecay}$ \Comment{Shrink the damping factor}\\ 
$\genstate_i^{n} \leftarrow  \genstate_{i-1}^{n}$ \Comment{Reset state to value at previous iteration}
}
}
Return converged solution, $\genstateArgnt{n}= \genstateArgnt{n}_i$
\end{algorithm}
\subsubsection{Numerical Considerations for the Forward and Backwards Problems}
The FBSM requires solving~\eqref{eq:lspg_continuous} and~\eqref{eq:lspg_adjoint}, both of which are defined at the time-continuous level. 
The numerical implementation of the FBSM requires two main ingredients: (1.) temporal discretization schemes for the forward and backward problems and (2.) 
an efficient method for computing action of the Jacobian transpose on a vector.
